---
title: "Lorenz, Pascal, 17-705-187, Group 14, Exercise 11"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Task 5
---

##a
```{r}
data <- read.table("http://stat.ethz.ch/Teaching/Datasets/banknoten.dat",header=TRUE)
real <- data$Laenge[1:100]
fake <- data$Laenge[101:200] #just to make my life a little easier instead of typing it out every time
t.test(real, mu=215)
t.test(fake, mu=215) #I use the default two-sided test, becaue we are testing for general differences of the mean, not specifically higher or lower values
```
These results show in multiple ways that the money marked as fake is actually fake:
1. The P-value for the mean of the sample of real value is very high (with 0-hypothesis of it being real money with mean length 215), while the P-value for the fake one is low (below 1%) with the same H0.
2. The true mean of 215 is contained within the 95 percent confidence interval, while the interval for the fake money does not overlap 215.


##b
```{r}
plot(density(fake),type="l",col="red",main="Distribution",xlab="Lenght")
lines(density(real), col="green")
abline(v=mean(real), col="green")
abline(v=mean(fake), col="red")
legend("topright",c("Real","Fake"), col=c("green","red"),pch=16)
```


##c
```{r}
t.test(real,fake, var.equal=TRUE) #var.equal=TRUE because of the task, though the difference is minimal in this case anyways (variance of real and fake money is very close)
```
